---
title: "Filterable plotly ggplots"
author: "Alex Zajichek"
date: "2023-08-09"
output: html_document
---

Here we go through a demonstration of how to build an HTML widget consisting of a `plotly` graph that is defined through `ggplot2` and has filters attached so the user can toggle through what to see in the graph. This is all self-contained in the resulting HTML output file. In short, we can have an interactive graph in our HTML document.

First, we need to load some packages.

```{r}
library(tidyverse)
library(survival)
library(plotly)
library(crosstalk)
```

Then, we'll construct a data set. Our focus will be on displaying survival curves for different subsets of patients.

```{r}
dat <- 
    survival::veteran |> 
    as_tibble() |>
    transmute(
        Treatment = paste0("Treatment ", trt),
        CellType = celltype,
        Age = 
            case_when(
                age > 60 ~ "Greater than 60",
                TRUE ~ "Less than (or equal) to 60"
            ),
        Time = time,
        Status = status
    )
dat
```

Next, we'll `nest` the data and extract the survival function from each subset.

```{r}
dat <-
    dat |>
    
    # Make a nested data set
    group_by(
        CellType,
        Age,
        Treatment
    ) |> 
    nest() |>
    
    # Get survival curves
    mutate(
        Survival = 
            data |> 
            map(
                function(.x) {
                    
                    # 1. Get the survival curve
                    temp_mod <- survfit(Surv(Time, Status) ~ 1, data = .x)
                    
                    # 2. Make a data frame of desired elements
                    tibble(
                        Time = temp_mod$time,
                        Estimate = temp_mod$surv,
                        Lower = temp_mod$lower,
                        Upper = temp_mod$upper,
                        AtRisk = temp_mod$n.risk
                    )
                    
                }
            )
    ) |>
    
    # Remove the data set
    select(-data) |> 
    
    # Expand the set 
    unnest(cols = Survival) |>
    ungroup()

# Make a shared data set
sd <-SharedData$new(dat)

# Make the cell type filter
filter_celltype <- 
    filter_select(
        id = "celltype",
        label = "CellType",
        sharedData = sd,
        group = ~CellType,
        multiple = FALSE
    )

# Make the age filter
filter_age <- 
    filter_select(
        id = "age",
        label = "Age",
        sharedData = sd,
        group = ~Age,
        multiple = FALSE
    )

# Put all UI objects into a container
bscols(
    
    # Put everything in a single columns
    shiny::fluidRow(
        shiny::column(
            width = 12,
            
            # Add a row of filters
            shiny::fluidRow(
                shiny::column(filter_celltype, width = 6),
                shiny::column(filter_age, width = 6)
            ),
            
            # Add the plot
            ggplotly(
                sd |>
                    
                    ggplot(
                        aes(
                            x = Time,
                            y = Estimate,
                            group = Treatment,
                            text = 
                                paste0(
                                    "<br>", Treatment,
                                    "<br>Time: ", round(Time, 2),
                                    "<br>Risk: ", round(Estimate*100, 1), "%",
                                    "<br>Lower: ", round(Lower*100, 1), "%",
                                    "<br>Upper: ", round(Upper*100, 1), "%",
                                    "<br>At-Risk: ", AtRisk
                                )
                        )
                    ) +
                    geom_line(
                        aes(
                            linetype = Treatment,
                            color = Treatment
                        ),
                        linewidth = 1
                    ) +
                    geom_ribbon(
                        aes(
                            ymin = Lower,
                            ymax = Upper,
                            fill = Treatment
                        ),
                        alpha = .25
                    ) + 
                    xlab("Time") +
                    scale_y_continuous(
                        name = "Survival probability (%)",
                        labels = scales::percent
                    ) + 
                    theme(
                        panel.background = element_blank(),
                        panel.grid.major.y = element_line(colour = "gray"),
                        panel.grid.major.x = element_blank(),
                        axis.title = element_text(size = 12),
                        axis.text = element_text(size = 12)
                    ) +
                    guides(fill = "none"),
                tooltip = "text"
            ) |>
                layout(legend = list(x = 0, y = 1))
        )
    )
)
```

Now, in order to make the filters display a selection by default, we can add some `JavaScript`. 

```{js}
function filter_default() {
    document.getElementById("celltype").getElementsByClassName("selectized") 
[0].selectize.setValue("smallcell", false);
 }
 function filter_default2() {
    document.getElementById("age").getElementsByClassName("selectized") 
[0].selectize.setValue("Greater than 60", false);
 }
window.addEventListener("load", filter_default)
window.addEventListener("load", filter_default2)
```

The `addEventListener` method tells these functions to run when the browser is loaded.